---
title: "Geo-Engineering Pilot Study using NetCDF (NC) Data"
output: html_document
date: "2024-06-10"
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE, error=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is an R/RStudio Pilot Project using [netCDF (.nc) data](https://www.unidata.ucar.edu/software/netcdf/) to explore Geo-Engineering ideas.

# Background

 - Check some of [these Geo-engineering Article Sources](https://www.economist.com/search?q=geoengineering).
 - Read [this article](https://www.economist.com/the-economist-explains/2024/04/16/what-is-geoengineering), which defines **Geo-engineering** as “*the deliberate large-scale manipulation of the planetary environment to counteract anthropogenic climate change*”. 


There a two complementary Geo-Engineering Strategies: 

 - those that make it easier for the Earth to *lose heat*, and 
 - those that *reduce the amount of heat gain*.

# Datasets

We can use [University of Washington Marine Cloud Brightening (MCB) programme data])(https://zenodo.org/records/8430485). Figure 4 in [this paper](https://gmd.copernicus.org/articles/17/2583/2024/) is generated by these data.
Additional data is available from the [Climate Research Unit](http://www.cru.uea.ac.uk/data).

 - [Download the *gmd_ukesm.tar.gz* file from this website](https://zenodo.org/records/8430485).
 - Use the `R` script below to load one of the many files in this tar-ball:

The examples make use of a netCDF file of climate data consisting of **predicted** long-term mean values (2033-02-14 through 2069-07-05, average of $720$ hours between $444$ elements) of near-surface air temperature on a grid of land points (longitude and latitude). The dimensions of the array are $192 (longitudes) \times 144 (latitudes) \times 444 (hours)$, thus forming a raster “stack” or “brick” consisting of $444$ layers.

```{r dataImport, warning=FALSE, message=FALSE, error=FALSE}
# install.packages("ncdf4", "raster", "archive")
library("ncdf4")
library("raster")
library(archive)
library(CFtime)
library(lattice)
library(RColorBrewer)

ncFile <- "C:/Users/IvoD/Desktop/gmd_ukesm.tar.gz"
fileList <- untar(ncFile, list=T)

ncin <- nc_open("C:/Users/IvoD/Desktop/ukesm_arise_ts_0NS.nc")
# ncin <- nc_open(archive_read(ncFile, file = "ukesm_ssp_ts_ave.nc"))
print(ncin)

# Longitude, in degrees
lon <- ncvar_get(ncin, "longitude")
nlon <- dim(lon)   # 192
head(lon)
# Latitude, in degrees
lat <- ncvar_get(ncin, "latitude")
nlat <- dim(lat)   # 192
head(lat)
# Time
time <- ncvar_get(ncin,"time")
length(time)
# Time units
tunits <- ncatt_get(ncin,"time","units")
tunits

# get temperature
dname <- "air_temperature"
tmp_array <- ncvar_get(ncin, dname)
dlname <- ncatt_get(ncin, dname, "long_name")
dunits <- ncatt_get(ncin, dname, "units")
fillvalue <- ncatt_get(ncin, dname, "_FillValue")
dim(tmp_array)  # [1] 192 144 444
str(tmp_array) # num [1:192, 1:144, 1:444] 

# get global attributes
# title <- ncatt_get(ncin, 0, "title")
# institution <- ncatt_get(ncin,0,"institution")
# datasource <- ncatt_get(ncin,0,"source")
# references <- ncatt_get(ncin,0,"references")
# history <- ncatt_get(ncin,0,"history")
# Conventions <- ncatt_get(ncin,0,"Conventions")

# convert time to CFtime class
cf <- CFtime(tunits$value, calendar = "proleptic_gregorian", time)
cf
timestamps <- CFtimestamp(cf) # get character-string times
# timestamps

time_cf <- CFparse(cf, timestamps) # parse the string into date components
# time_cf
```

`r fileList`

# Some  Plots


```{r graphExamples, echo=TRUE, fig.height=12, fig.width=12, warning=FALSE, message=FALSE, error=FALSE}
library(plotly)
library(data.table)

# get a single slice or layer (January)
m1 <- 10 # "2033-11-11" # the 10-th element
tmp_slice <- tmp_array[ , , m1]

# image(lon, lat, tmp_slice, col=rev(brewer.pal(10,"RdBu")))
df1 <- data.table(longitude = rep(lon, ncol(tmp_slice)), 
                 latitude = rep(lat, each = nrow(tmp_slice)),
                 value = c(tmp_slice))

# Surface 2D projection Map
# g <- list(showland = TRUE, landcolor = toRGB("#e5ecf6"))
# fig <- plot_ly(type = 'scattergeo', mode = 'markers')
# fig <- fig %>% layout(geo = g)
# fig

# 3D Dynamic SUrface Map
g1 <- list(projection = list(type = 'orthographic'), showland = TRUE, 
          landcolor = "lightgray", lakecolor = "lightblue") # toRGB("#e5ecf6")
fig2033 <- plot_ly(type = 'scattergeo', mode = 'markers') %>% 
  add_trace(x=df1$longitude, y=df1$latitude, color=df1$value, 
            opacity=0.8, size=df1$value, showlegend=F) %>% 
  layout(geo = g1, title="(G6-1.5) A Geoengineering Model Intercomparison Project \n (GeoMIP): Predicted Sunlight Reflection (2033-11-11)") %>% hide_colorbar()
fig2033

# tail(time_cf) # 444 2069     7   5 
m2 <- 444 # "2069-07-5" # the 444-th element
tmp_slice <- tmp_array[ , , m2]

# image(lon, lat, tmp_slice, col=rev(brewer.pal(10,"RdBu")))
df2 <- data.table(longitude = rep(lon, ncol(tmp_slice)), 
                 latitude = rep(lat, each = nrow(tmp_slice)),
                 value = c(tmp_slice))

# Surface 2D projection Map
# g <- list(showland = TRUE, landcolor = toRGB("#e5ecf6"))
# fig <- plot_ly(type = 'scattergeo', mode = 'markers')
# fig <- fig %>% layout(geo = g)
# fig

# 3D Dynamic SUrface Map
g2 <- list(projection = list(type = 'orthographic'), showland = TRUE, 
          landcolor = "lightgray", lakecolor = "lightblue") # toRGB("#e5ecf6")
fig2069 <- plot_ly(type = 'scattergeo', mode = 'markers') %>% 
  add_trace(x=df2$longitude, y=df2$latitude, color=df2$value, 
            opacity=0.8, showlegend=F) %>% 
  layout(geo = g2, title="(G6-1.5) A Geoengineering Model Intercomparison Project \n (GeoMIP): Predicted Sunlight Reflection (2069-07-5)") %>% hide_colorbar()
fig2069

g3 <- list(projection = list(type = 'orthographic'), showland = TRUE, 
          landcolor = "lightgray", lakecolor = "lightblue") # toRGB("#e5ecf6")
diff_fig <- plot_ly(type = 'scattergeo', mode = 'markers') %>% 
  add_trace(x=df2$longitude, y=df2$latitude, color=(df1$value- df2$value), 
            opacity=0.8, showlegend=F) %>% 
  layout(geo = g3, title="(G6-1.5) A Geoengineering Model Intercomparison Project \n (GeoMIP): Predicted Diff Sunlight Reflection (2033 vs. 2069)") %>% hide_colorbar()
diff_fig

# ... Create the Synched 2033 and 2069 Models, and their Difference ...
# Combine them into a subplot
# main_plot <- subplot(fig2033, fig2069, diff_fig) %>%
#   layout(
#     title = "(G6-1.5) A Geoengineering Model Intercomparison Project \n (GeoMIP): Predicted Diff Sunlight Reflection (2033 vs. 2069)",
#     scene  = list(domain = list(x = c(0, 0.5), y = c(0.5, 1)), aspectmode = "cube"),
#     scene2 = list(domain = list(x = c(0.5, 1), y = c(0.5, 1)), aspectmode = "cube"),
#     scene3 = list(domain = list(x = c(0.25, 0.75), y = c(0, 0.5)), aspectmode = "cube")
#   )
# #https://stackoverflow.com/questions/77757102/possible-to-synchronize-3d-scatter-plotly-plots-in-r
# 
# main_plot %>%
#   htmlwidgets::onRender(
#     "function(x, el) {
#       x.on('plotly_relayout', function(d) {
#         const camera = Object.keys(d).filter((key) => /\\.camera$/.test(key));
#         if (camera.length) {
#           const scenes = Object.keys(x.layout).filter((key) => /^scene\\d*/.test(key));
#           const new_layout = {};
#           scenes.forEach(key => {
#             new_layout[key] = {...x.layout[key], camera: {...d[camera]}};
#           });
#           Plotly.relayout(x, new_layout);
#         }
#       });
#     }")
```

...

